Verifying Cache Coherence Protocols for an Infinite Family of Cache Hierarchies

### 2014-02-06 paper/Conclusion.tex
Prune related work
Pass over Conclusion and some capitalization of other section titles

\section{Conclusions and future work}

In this paper, we provided a formal definition of cache coherence via the store
atomicity property and provided a concrete hierarchy-parametric directory-based
MSI protocol whose correctness we formally checked via a Coq-based proof. To our
knowledge this is the first instance of a formal proof of a protocol parametric
over the cache hierarchy. We have also provided a set of invariants not
specific to MSI which imply store atomicity; other protocols that obey these
invariants are already proven to be correct. Finally, we have identified a set
of key properties that seem to be shared with many other protocols which ease
the verification considerably.

Our protocol is described in a style amenable to mechanical transliteration to a
Bluespec SystemVerilog hardware description allowing us to provide high-quality
verified cache coherence engines. As the protocol and associated proof are robust
to many microarchitecture-oriented changes (\eg{} arbitrary hierarchy and
arbitrary cache size, associativity, replacement policy choices, policies to
handle incoming requests, \etc{} in each node), we should be able to produce a
high performance performance formally verified hardware cache coherence system.

Moving this work towards generating a fully verified MSI implementation
requires a number of small steps. First, the simple transliteration of
descriptions between Bluespec and Coq, must be done in a mechanical way.
Second, the compilation of Bluespec to hardware is unverified; to fully trust
the implementation the Bluespec compiler must be verified. Some initial work on
this has already been done~\cite{TDBLP:conf/cav/BraibantC13}. Finally, we must
extend our notion of correctness to provide some guarantee of
liveness and fairness in handling of requests from processors. We have
initial Coq-based proofs that with finite buffers a local fairness in
transaction our protocol will have this property, but more work is needed.

>>>>

\section{Conclusions and Future Work}

In this paper, we provided a formal definition of cache coherence via the store
atomicity property and provided a concrete hierarchy-parametric directory-based
MSI protocol whose correctness we formally checked via a Coq-based proof. To our
knowledge, this is the first instance of formal verification of a protocol parametric
over the cache hierarchy. We have also provided a set of invariants not
specific to MSI that imply store atomicity; other protocols that obey these
invariants are already proven to be correct. Finally, we have identified a set
of key properties that seem to be shared with many other protocols and which ease
the verification considerably.

Our protocol is described in a style amenable to mechanical transliteration to a
Bluespec SystemVerilog hardware description, allowing us to provide high-quality
verified cache coherence engines. As the protocol and associated proof are robust
to many microarchitecture-oriented changes (\eg{} arbitrary hierarchy and
arbitrary cache size, associativity, replacement policy choices, policies to
handle incoming requests, \etc{} in each node), we should be able to produce a
high-performance formally verified hardware cache coherence system.

Moving this work towards generating a fully verified MSI implementation
requires a number of small steps. First, the simple transliteration of
descriptions between Bluespec and Coq must be done in a mechanical way.
Second, the compilation of Bluespec to hardware is unverified; to fully trust
the implementation the Bluespec compiler must be verified. Some first steps
have been taken in that direction~\cite{TDBLP:conf/cav/BraibantC13}. Finally, we must
extend our notion of correctness to provide some guarantee of
liveness and fairness in handling of requests from processors. We have
initial Coq-based proofs that with finite buffers a local fairness condition
holds for our protocol, but more work is needed.
